Coherent optical creation of a single molecule

### 2020-12-09 paper.tex
Update on Overleaf.

\begin{figure*}
  \includegraphics[width=\textwidth]{imgs/fig-theory.pdf}
  \caption{Optical creation of single molecules from single atoms in an optical tweezer.
    (a) Schematic of the optical transition from an atom pair to a weakly bound molecule.
    The initial state is the relative motional ground state between the two atoms
    and the final state is the first molecular bound state.
    The transition is driven by a pair of laser frequencies whose difference ($< 1$~GHz) matches the molecular binding energy.
    The lasers are detuned by $\Delta$ from an excited molecular state
    in the $\mathrm{c^3\Sigma}$ electronic state
    in order to reduce scattering during the transfer.
    (b) Calculation of the scattering rate vs. laser frequency
    including all vibrational states of 8 excited-state potentials
    and the atomic continuum.
    The assumed excited-state linewidth for molecular lines is $50~\text{MHz}$.
    The dotted red line shows the contribution from the near-threshold states, which approximately scales as $1/\Delta_{\mathrm{thresh}}^2$.
    (c) Comparison between vibrational sublevels of the intermediate excited state
    for the Raman transition.
    The lowest sublevel (blue)
    has a larger optimum ratio of Raman Rabi frequency to scattering rate
    than the higher sublevel (orange).
    (d) The large scattering length for $\Na(2,2),\Cs(3,3)$ is associated with enhancementment of the relative wavefunction at short
    internuclear distance ($R$).
    \label{f-theory}
  }
\end{figure*}

@@

The optical Raman transfer is illustrated
by the idealized three-level system shown in Fig.~\ref{f-theory}a,
where the initial atomic state and the target weakly bound molecular state are coupled to an intermediate state by two lasers with Rabi frequencies $\Omega_{\rm a}$ and $\Omega_{\rm m}$, one-photon detuning $ \Delta $, and all Rabi frequencies are population oscillation frequencies.  %(Ultimately, we calculate with contribution of all vibrational states of the excited electronic potential, $\mathrm{c^3\Sigma^+}$)
The transfer Raman Rabi frequency is given by $\Omega_{\rm a}\Omega_{\rm m} / (2\Delta)$~\cite{Wineland2003}.
Unlike Raman transitions in atoms, the one-photon Rabi frequencies are greatly imbalanced ($\Omega_{\rm a}/\Omega_{\rm m} \ll 1$)
due to the small wavefunction overlap between the atomic state and the intermediate state,
and scattering losses are dominated by the final state. Furthermore, the energy difference between the atomic state and target molecular state is small ($ < 1~\mathrm{GHz} $) compared to the single-photon detuning of $80$ to $200$~GHz, causing the target molecular state to scatter both beams nearly equally with a total rate $ \Gamma_{\rm e} \Omega_{\rm m}^2 / (2\Delta^2)$, where $ \Gamma_{\rm e} $ is the excited-state linewidth~\footnote{The two beams have equal power to maximize the Raman Rabi frequency at fixed total power, resulting in a factor of 2 for scattering two beams.}.
In this idealized treatment, the ratio between the Raman Rabi frequency and the scattering rate is $ \Omega_{\rm a}/\Omega_{\rm m} \times \Delta/\Gamma_{\rm e} $, which limits the transfer efficiency into the molecular state. At the same time, the intensity-stability requirement is determined by the ratio of Raman Rabi frequency to light shift ($\Omega_{\rm a}/\Omega_{\rm m}$)~\footnote{There is an additional factor of 2, with both beams at equal power,
  to account for the Stark shift caused by both beams.}. Notably, both figures of merit improve with a larger ratio $\Omega_{\rm a}/\Omega_{\rm m}$. %, depends on the ratio of the two matrix elements and how far detuned the laser is from the transition in units of the linewidth.
% However, the detuning cannot be too large, since that will reduce the Raman Rabi frequency.

@@

\begin{figure}[t!]
  \includegraphics[width=0.48\textwidth]{imgs/fig-raman.pdf}
  \caption{Coherent transfer of atoms to molecules.  The molecular state is dark to the imaging step and corresponds to zero signal.
    (a) Raman difference frequency scans for various durations
    showing the resonance as the detuning from
    the fitted resonance at $770.5715(1)$~MHz, near the calculated value of $763$~MHz.
    (b) Raman pulse-length scan on resonance.
    A decaying Rabi oscillation shows the coherence of
    the Raman transfer process.
    A model is fitted to (a) and (b) to determine
    the Raman Rabi frequency and loss rates.
    \emph{Inset:} Geometry and polarization of trap and Raman beam relative to the magnetic field.
    The 3.25~mW beam is focused to a waist of $0.9~\mathrm{\mu m}$
    that confines the atoms and molecule.
    A perpendicular $B_0=8.83(2)~\mathrm{G}$ magnetic field
    defines the quantization axis and the atoms experience predominantly $\pi$-polarized light.
    \label{f-raman}}
\end{figure}
After merging the tweezers, we drive the atoms into spin combination $\ket{\uparrow_{\Na} \downarrow_{\Cs}}$ with a large scattering length
by performing a Cs spin flip while taking into account
the $-30.7~\mathrm{kHz}$ interaction shift~\cite{Hood2019}.
This is the initial atomic state for Raman transfer.
The spin flip selectively transfers atoms in the relative motional ground state,
removing any background from atoms in excited states of relative motion
\footnote{This interaction shift is larger than the differential axial trapping frequency
  between Na and Cs atoms, which decouples the relative and center of mass motional state
  and improves the robustness of our preparation of the relative motional ground state.}.
For the experiment reported here,
$31\mathrm{\%}$ of initial two-atom population is transferred. Of this population, over $60\%$ is in the ground state of center-of-mass motion, inferred from Raman sideband thermometry.

\begin{figure*}[htb]
  \includegraphics[width=\textwidth]{imgs/fig-det.pdf}
  \caption{Raman transition parameters as a function of tweezer power and detuning.
    (a) The Raman resonance~($\omega_{\rm R}$) fitted by $a_P+b_P/\Delta$, where
    $a_P$ and $b_P=(\Omega_{\rm a}^2-\Omega_{\rm m}^2)/2$
    are the power ($P$) dependent background and $v'=0$ contributions
    to the light shift and
    $\Delta\equiv2\pi\times\paren{f_0 - f_{\mathrm{tweezer}}}$ is the detuning from
    the $v'=0$ resonance frequency $f_0$.
    $a_P$ is fitted to a model including linear and small quadratic light shift
    \todo{which assumes $\Omega_{\rm m}\gg\Omega_{\rm a}$} to obtain the Raman resonance frequency
    at zero tweezer power $\omega_{\rm R0}=2\pi\times770.1969(2)~\mathrm{MHz}$ where the statistical uncertainty is shown.
    (b) Raman Rabi frequency ($\Omega_{\rm R}$) fitted to $c_P+d_P/\Delta$, where
    $c_P$ and $d_P=\Omega_{\rm a}\Omega_{\rm m}/2$
    are the background and $v'=0$ contributions that scale as $P^{1.29}$ with optical power.
    The detuning is calculated from the PA frequency fitted in (a). Fit parameters are listed in Table~\ref{tab:f-det:fit}.
    (c) Tweezer power dependency of $\Omega_{\rm m}$ (top) and $\Omega_{\rm a}$ (bottom) calculated from
    $b_P$ and $d_P$ on a log-log scale showing approximate $P^{0.5}$ scaling of $\Omega_{\rm m}$ and
    $P^{0.79}$ scaling of $\Omega_{\rm a}$.
    \label{f-det}}
\end{figure*}

@@

To understand the details and limitations of the Raman transfer process better,
we measured the properties of the two-photon resonance as a function of tweezer power and single-photon detuning.
Known dependencies of the light shift and Raman Rabi frequency ($\Omega_{\rm R}$) on detuning ($\Delta$) allow experimental determination of the Rabi frequencies
$ \Omega_{\rm a} $ and $\Omega_{\rm m}$ whose ratio critically affects the transfer efficiency.
Both the light shift and $\Omega_{\rm R}$ follow a $1/\Delta$ slope as shown in Fig.~\ref{f-det}a, b and include a constant offset that we attribute to coupling to other excited states that are further away in energy.
The $1/\Delta$ components due to the nearby $v'=0$ intermediate state determine $\Omega_{\rm m} $ and $ \Omega_{\rm a} $ in Table~\ref{tab:f-det:fit}.

\begin{table}[ht]
  \centering
  \begin{tabular}{|c|c|c|c|}
    $P~(\mathrm{mW})$&$0.75$&$1.5$&$3.75$\\\hline
    $f_{\mathrm{PA}0}~(\mathrm{GHz})$&\multicolumn{3}{|c|}{$288711.8$}\\\hline
    $a~(\mathrm{2\pi\times MHz})$&$770.20452(6)$&$770.2081(1)$&$770.1943(3)$\\
    $b~(\mathrm{4\pi^2\times MHz\cdot GHz})$&$-12.46(2)$&$-24.44(3)$&$-60.66(8)$\\\hline
    $c~(\mathrm{2\pi\times kHz})$&$0.29(2)$&$0.63(4)$&$2.4(2)$\\
    $d~(\mathrm{4\pi^2\times MHz\cdot GHz})$&$0.115(4)$&$0.275(6)$&$0.95(3)$\\ \hline
    $\Omega_{\rm R}~(\mathrm{2\pi\times kHz})$ & 0.49(2) & 1.18(3) & 3.97(9) \\
    $\Omega_{\rm m}~(\mathrm{2\pi\times MHz})$ & & & 348.3(3) \\
    $\Omega_{\rm a}~(\mathrm{2\pi\times kHz})$ & & & 5.5(2)
  \end{tabular}
  \caption{Fitting results for Fig.~\ref{f-det}(a,b). $\Omega_{\rm R}$ is reported at $-151~\mathrm{GHz}$ detuning from the $v' = 0$ state.  The coefficients $\Omega_{\rm m}$ and $\Omega_{\rm a}$ are in broad agreement with our calculation and their ratio at $3.75~\mathrm{mW}$ is $0.016$, near the theory prediction of $0.013$.  The measured Rabi rate $\Omega_{\rm R}$ is only 63~\% of $\Omega_{\rm m} \Omega_{\rm a}/(2\Delta)$ due to interference from further-detuned Raman processes.
    \label{tab:f-det:fit}}
\end{table}

@@

The molecular lifetime of $0.199(9)~\mathrm{ms}$ in Fig.~\ref{f-lifetime}a is measured by preparing the molecule with a $\pi$-pulse,
followed after a variable delay by a second dissociating $\pi$-pulse.
This result is
consistent with the decay of the Rabi oscillation in Fig.~\ref{f-raman}b. Preliminary experiments and theoretical considerations indicate that the molecular lifetime may be limited by two-photon coupling to the atomic continuum~\cite{YichaoYu}.  Atom loss is shown to be small in Fig.~\ref{f-lifetime}b by measuring the two-body lifetimes of the atoms directly without the second Raman frequency. In principle, destructive interference that reduces the Raman Rabi rate ($\Omega_{\rm R}$) for negative detunings ($\Delta$) changes to constructive interference for positive detunings, but additional molecular resonances make the positive region unusable.  More negative detunings that might reduce the scattering rate were prevented by vanishing of $\Omega_{\rm R}$.

Separately we observe a decrease in coherence by a factor of 2 without laser spectrum filters,
suggesting that spectral impurity of the laser can be a significant source of loss.   While we have not fully characterized the sources
of broadband noise, possibilities include amplified spontaneous emission (ASE) from the laser and fiber nonlinearities. Other potential decoherence sources include
fluctuations of the tweezer intensity and magnetic field, although the shape of the Rabi oscillation in Fig.~{f-raman}b indicates that loss rather than frequency fluctuations are responsible for decoherence.
Based on the ratio between $\Omega_{\rm R}$ and the light shift,
the requirement on the tweezer intensity stability is $1\mathrm{\%}$ at $3.75~\mathrm{mW}$ power. We stabilize the power to $0.1\mathrm{\%}$, indicating that in the absence of beam-waist fluctuations, light shift is not a major source of decoherence.
Similarly, the measured Zeeman shift of $42.2(2)~\mathrm{kHz/G}$
does not cause significant decoherence for the measured magnetic field
fluctuation of $1.5~\mathrm{mG}$.

@@

\TR{Freq./mechanism of leakage light in Fig. 2?}

\TR{Insert lower practical limit for $\Delta$ as ``XXX GHz'' on last page.}

>>>>

\begin{figure*}
  \includegraphics[width=\textwidth]{imgs/fig-theory.pdf}
  \caption{Optical creation of single molecules from single atoms in an optical tweezer.
    (a) Schematic of the optical transition from an atom pair to a weakly bound molecule.
    The initial state is the relative motional ground state between the two atoms
    and the final state is the first molecular bound state.
    The transition is driven by a pair of laser frequencies whose difference ($< 1$~GHz) matches the molecular binding energy.
    The lasers are detuned by $\Delta$ from an excited vibrational sublevel
    in the $\mathrm{c^3\Sigma^+(\Omega = 1)}$ electronic state
    in order to reduce scattering during the transfer.
    (b) Calculation of the scattering rate vs. laser frequency
    including all vibrational states of 8 excited-state potentials
    and the atomic continuum.
    The assumed excited-state linewidth for molecular lines is $50~\text{MHz}$.
    The dotted red line shows the contribution from the near-threshold states, which approximately scales as $1/\Delta_{\mathrm{thresh}}^2$.
    (c) Comparison between vibrational sublevels of the intermediate excited state
    for the Raman transition.
    The lowest sublevel (blue)
    has a larger optimum ratio of Raman Rabi frequency to scattering rate
    than the higher sublevel (orange).
    (d) The large scattering length for $\Na(2,2),\Cs(3,3)$ is associated with enhancement of the relative wavefunction at short
    internuclear distance ($R$).
    \label{f-theory}
  }
\end{figure*}

@@

The optical Raman transfer is illustrated
by the idealized three-level system shown in Fig.~\ref{f-theory}a,
where the initial atomic state and the target weakly bound molecular state are coupled to an intermediate state by two lasers with Rabi frequencies $\Omega_{\rm a}$ and $\Omega_{\rm m}$, one-photon detuning $ \Delta $, and all Rabi frequencies are population oscillation frequencies.  %(Ultimately, we calculate with contribution of all vibrational states of the excited electronic potential, $\mathrm{c^3\Sigma^+}$)
The transfer Raman Rabi frequency is given by $\Omega_{\rm a}\Omega_{\rm m} / (2\Delta)$~\cite{Wineland2003}.
Unlike Raman transitions in atoms, the one-photon Rabi frequencies are greatly imbalanced ($\Omega_{\rm a}/\Omega_{\rm m} \ll 1$)
due to the small wavefunction overlap between the atomic state and the intermediate state,
and scattering losses are dominated by the final state. Because the energy difference between the atomic state and target molecular state is small ($ < 1~\mathrm{GHz} $) compared to the single-photon detuning of $80$ to $200$~GHz, both beams scatter nearly equally with a total rate $ \Gamma_{\rm e} \Omega_{\rm m}^2 / (2\Delta^2)$, where $ \Gamma_{\rm e} $ is the excited-state linewidth.  The two beams have equal power to maximize the Raman Rabi frequency at fixed total power.
In this idealized treatment, the ratio between the Raman Rabi frequency and the scattering rate is $ \Omega_{\rm a}/\Omega_{\rm m} \times \Delta/\Gamma_{\rm e} $, which limits the transfer efficiency into the molecular state. At the same time, the intensity-stability requirement is determined by the ratio of Raman Rabi frequency to light shift $\Omega_{\rm a}/\Omega_{\rm m}$. Notably, both figures of merit improve with a larger ratio $\Omega_{\rm a}/\Omega_{\rm m}$. %, depends on the ratio of the two matrix elements and how far detuned the laser is from the transition in units of the linewidth.
% However, the detuning cannot be too large, since that will reduce the Raman Rabi frequency.

@@

\begin{figure}[t!]
  \includegraphics[width=0.48\textwidth]{imgs/fig-raman.pdf}
  \caption{Coherent transfer of atoms to molecules.  The molecular state is dark to the imaging step and corresponds to zero signal.
    (a) Raman difference frequency scans for various durations
    showing the resonance as a function of the detuning from
    the fitted resonance at $770.5715(1)$~MHz, near the calculated value of $763$~MHz.
    (b) Raman pulse-length scan on resonance.
    A decaying Rabi oscillation shows the coherence of
    the Raman transfer process.
    A model is fitted to (a) and (b) to determine
    the Raman Rabi frequency and loss rates.
    \emph{Inset:} Geometry and polarization of trap and Raman beam relative to the magnetic field.
    The 3.25~mW beam is focused to a waist of $0.9~\mathrm{\mu m}$
    that confines the atoms and molecule.
    A perpendicular $B_0=8.83(2)~\mathrm{G}$ magnetic field
    defines the quantization axis and the atoms experience predominantly $\pi$-polarized light.
    \label{f-raman}}
\end{figure}
After merging the tweezers, we drive the atoms into spin combination $\ket{\uparrow_{\Na} \downarrow_{\Cs}}$ with a large scattering length
by performing a Cs spin flip while taking into account
the $-30.7~\mathrm{kHz}$ interaction shift~\cite{Hood2019}.
This is the initial atomic state for Raman transfer.
The spin flip selectively transfers atoms in the relative motional ground state,
removing any background from atoms in excited states of relative motion
\footnote{This interaction shift is larger than the differential axial trapping frequency
  between Na and Cs atoms, which decouples the relative and center of mass motional state
  and improves the robustness of our preparation of the relative motional ground state.}.
For the experiment reported here,
$31\mathrm{\%}$ of initial two-atom population is transferred. Of this population, over $60\%$ is in the ground state of center-of-mass motion, inferred from Raman sideband thermometry.

\begin{figure*}[htb]
  \includegraphics[width=\textwidth]{imgs/fig-det.pdf}
  \caption{Raman transition parameters as a function of tweezer power and detuning.
    (a) The Raman resonance $\omega_{\rm R}$ fitted to $a_P+b_P/\Delta$, where
    $a_P$ and $b_P=(\Omega_{\rm a}^2-\Omega_{\rm m}^2)/2$
    are the power ($P$) dependent background and $v'=0$ contributions
    to the light shift and
    $\Delta\equiv2\pi\times\paren{f_{\mathrm{PA}0} - f_{\mathrm{tweezer}}}$ is the detuning from
    the $v'=0$ resonance frequency $f_{\mathrm{PA}0}$.
    $a_P$ is fitted to a model including linear and small quadratic light shift
    \todo{which assumes $\Omega_{\rm m}\gg\Omega_{\rm a}$} to obtain the Raman resonance frequency
    at zero tweezer power $\omega_{\rm R0}=2\pi\times770.1969(2)~\mathrm{MHz}$ where the statistical uncertainty is shown.
    (b) Raman Rabi frequency $\Omega_{\rm R}$ fitted to $c_P+d_P/\Delta$, where
    $c_P$ and $d_P=\Omega_{\rm a}\Omega_{\rm m}/2$
    are the background and $v'=0$ contributions that scale as $P^{1.29}$.
    The detuning is calculated from $f_{\mathrm{PA}0}$ fitted in (a). Fit parameters are listed in Table~\ref{tab:f-det:fit}.
    (c) Tweezer power dependency of $\Omega_{\rm m}$ (top) and $\Omega_{\rm a}$ (bottom) calculated from
    $b_P$ and $d_P$ on a log-log scale showing approximate $P^{0.5}$ scaling of $\Omega_{\rm m}$ and
    $P^{0.79}$ scaling of $\Omega_{\rm a}$.
    \label{f-det}}
\end{figure*}

@@

To understand the details and limitations of the Raman transfer process better,
we measured the properties of the two-photon resonance as a function of tweezer power and single-photon detuning.
Known dependencies of the light shift and Raman Rabi frequency $\Omega_{\rm R}$ on detuning $\Delta$ allow experimental determination of the Rabi frequencies
$ \Omega_{\rm a} $ and $\Omega_{\rm m}$ whose ratio critically affects the transfer efficiency.
Both the light shift and $\Omega_{\rm R}$ follow a $1/\Delta$ slope as shown in Fig.~\ref{f-det}a, b and include a constant offset that we attribute to coupling to other excited states that are further away in energy.
The $1/\Delta$ components due to the nearby $v'=0$ intermediate state determine $\Omega_{\rm m} $ and $ \Omega_{\rm a} $ in Table~\ref{tab:f-det:fit}.

\begin{table}[ht]
  \centering
  \begin{tabular}{|c|c|c|c|}
    $P~(\mathrm{mW})$&$0.75$&$1.5$&$3.75$\\\hline
    $f_{\mathrm{PA}0}~(\mathrm{GHz})$&\multicolumn{3}{|c|}{$288711.8$}\\\hline
    $a~(\mathrm{2\pi\times MHz})$&$770.20452(6)$&$770.2081(1)$&$770.1943(3)$\\
    $b~(\mathrm{4\pi^2\times MHz\cdot GHz})$&$-12.46(2)$&$-24.44(3)$&$-60.66(8)$\\\hline
    $c~(\mathrm{2\pi\times kHz})$&$0.29(2)$&$0.63(4)$&$2.4(2)$\\
    $d~(\mathrm{4\pi^2\times MHz\cdot GHz})$&$0.115(4)$&$0.275(6)$&$0.95(3)$\\ \hline
    $\Omega_{\rm R}~(\mathrm{2\pi\times kHz})$ & 0.49(2) & 1.18(3) & 3.97(9) \\
    $\Omega_{\rm m}~(\mathrm{2\pi\times MHz})$ & & & 348.3(3) \\
    $\Omega_{\rm a}~(\mathrm{2\pi\times kHz})$ & & & 5.5(2)
  \end{tabular}
  \caption{Fitting results for Fig.~\ref{f-det}(a,b). $\Omega_{\rm R}$ is reported at $-151~\mathrm{GHz}$ detuning from the $v' = 0$ state.  At $3.75~\mathrm{mW}$, $\Omega_{\rm m}/\Omega_{\rm a}=0.016$, near the theory prediction of $0.013$.  The measured Rabi rate $\Omega_{\rm R}$ is only 63~\% of $\Omega_{\rm m} \Omega_{\rm a}/(2\Delta)$ due to interference from further-detuned Raman processes.
    \label{tab:f-det:fit}}
\end{table}

@@

The molecular lifetime of $0.199(9)~\mathrm{ms}$ in Fig.~\ref{f-lifetime}a is measured by preparing the molecule with a $\pi$-pulse,
followed after a variable delay by a second dissociating $\pi$-pulse.
This result is
consistent with the decay of the Rabi oscillation in Fig.~\ref{f-raman}b. Preliminary experiments and theoretical considerations indicate that the molecular lifetime may be limited by two-photon coupling to the atomic continuum~\cite{YichaoYu}.  Atom loss is shown to be small in Fig.~\ref{f-lifetime}b by measuring the two-body lifetimes of the atoms directly without the second Raman frequency. In principle, destructive interference that reduces the Raman Rabi rate $\Omega_{\rm R}$ for negative detunings $\Delta$ changes to constructive interference for positive detunings, but additional molecular resonances make the positive region unusable.  More negative detunings that might reduce the scattering rate were prevented by vanishing of $\Omega_{\rm R}$.

Separately we observe a decrease in coherence by a factor of 2 without laser spectrum filters,
suggesting that spectral impurity of the laser can be a significant source of loss. While we have not fully characterized the sources
of broadband noise, possibilities include amplified spontaneous emission (ASE) from the laser and fiber nonlinearities. Other potential decoherence sources include
fluctuations of the tweezer intensity and magnetic field, although the shape of the Rabi oscillation in Fig.~\ref{f-raman}b indicates that loss rather than frequency fluctuations are responsible for decoherence.
Based on the ratio between $\Omega_{\rm R}$ and the light shift,
the requirement on the tweezer intensity stability is $1\mathrm{\%}$ at $3.75~\mathrm{mW}$ power. We stabilize the power to $0.1\mathrm{\%}$, indicating that in the absence of beam-waist fluctuations, light shift is not a major source of decoherence.
Similarly, the measured Zeeman shift of $42.2(2)~\mathrm{kHz/G}$
does not cause significant decoherence for the measured magnetic field
fluctuation of $1.5~\mathrm{mG}$.

@@

\TR{What is meant by ``result is consistent with the decay of the Rabi oscillation'' if independent fits not done?  See also Fig.~5 caption.}
